---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
  collapse = F,
  comment = "",
  fig.path = "README-",
  warning = F,
  message = F,
  dpi=300,
  dev = "png"
)
library(RSDA)
```

# RSDA

Symbolic Data Analysis (SDA) was proposed by professor Edwin Diday in 1987, the main purpose of SDA is to substitute the set of rows (cases) in the data table for  a concept (second order statistical unit). This package implements, to the symbolic case, certain techniques of automatic classification, as well as some linear models.

## Installation

You can install RSDA from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("PROMiDAT/RSDA")
```

## Examples

```{r example}
data(ex1_db2so)
ex1_db2so
```

```{r,warning=FALSE,message=FALSE}
result1 <- classic.to.sym(ex1_db2so, concept=c("state"),
                    variables=c("county", "group","sex", "age","age"),
                    variables.types=c("$S", "$M","$M","$I","$C"))
result1

result2 <- classic.to.sym(ex1_db2so, concept=c("state", "sex"),
                    variables=c("county", "group", "age"),
                    variables.types=c("$S", "$M", "$I"))
result2

result3 <- classic.to.sym(ex1_db2so, concept=c("state"),
                    variables=c("age"),
                    variables.types=c("$H"))
result3
```

# Basic plots in RSDA

```{r,fig.align='center'}
data(oils)
oils
sym.scatterplot(oils[,3], oils[,4], main='Scatter Plot')
sym.radar.plot(oils[1:2,],main = "Radar Plot")
sym.radar.plot(oils[1:4,],main = "Radar Plot")
h <- interval.histogram.plot(oils[,3],n.bins = 2,main = "histogram")
h <- interval.histogram.plot(oils[,2],n.bins = 3,main = "histogram", col = "blue")
```


# Basic statistics in RSDA

## Mean

```{r}
result1
mean(result1[,5])
mean(result1[,4],method = "interval")
mean(result1[,3],method = "modal")
```


## Standard Deviation

```{r}
sd(result1[,5])
sd(result1[,4],method = "interval")
sd(result1[,3],method = "modal")
```

## Varianze

```{r}
var(result1[,5])
var(result1[,4],method = "interval")
var(result1[,3],method = "modal")
```

## Correlation

```{r}
cor(oils[,1],oils[,2])
cor(oils[,2],oils[,3],method = "billard")
cor(oils[,3],oils[,4])
```

# Principal Component Analysis for Interval Data

```{r}
oils
```

## Example 1
```{r,fig.align='center'}
res <- sym.interval.pca(oils,'centers')
sym.scatterplot(res$Sym.Components[,1], res$Sym.Components[,2],
                 labels=TRUE,col='red',main='Centers PCA Method - Oils Data')
sym.circle.plot(res$Sym.Prin.Correlations)
```

## Example 2
```{r,fig.align='center'}
res <- sym.interval.pca(oils,'tops')
sym.scatterplot(res$Sym.Components[,1],res$Sym.Components[,2],
                labels=TRUE,col='red',main='PCA Vertex - Oil Data')
```

## Example 3

```{r,fig.align='center'}
res <- sym.interval.pca(oils,'principal.curves')
sym.scatterplot(res$sym.prin.curve[,1],res$sym.prin.curve[,2],
                 labels=TRUE,col='red',main='Principal Curves PCA - Oils Data')
```

## Example 4

```{r,fig.align='center'}
res <- sym.interval.pca(oils,'optimized.distance')
sym.scatterplot(res$Sym.Components[,1] , res$Sym.Components[,2],
labels = TRUE,col='red',main='Optimized PCA Distance - Oils Data')
plot(res$pca.min,choix = "var")
```

## Example 5

```{r,fig.align='center'}
res <- sym.interval.pca(oils,'optimized.variance')
sym.scatterplot(res$Sym.Components[,1] ,res$Sym.Components[,2],
labels = TRUE,col='red',main='Optimized PCA Variance - Oils Data')
plot(res$pca.min,choix = "var")
```

# Regression for Interval Data

## Lineal Model

```{r}
data(int_prost_train)
data(int_prost_test)
res.cm<-sym.lm(lpsa~.,sym.data=int_prost_train,method='cm')
pred.cm<-predictsym.lm(res.cm,int_prost_test,method='cm')
RMSE.L(sym.var(int_prost_test,9),pred.cm$Fitted)
RMSE.U(sym.var(int_prost_test,9),pred.cm$Fitted)
R2.L(sym.var(int_prost_test,9),pred.cm$Fitted)
R2.U(sym.var(int_prost_test,9),pred.cm$Fitted)
deter.coefficient(sym.var(int_prost_test,9),pred.cm$Fitted)
```

## LASSO

```{r,fig.align='center'}
data(int_prost_train)
data(int_prost_test)
res.cm.lasso<-sym.glm(sym.data=int_prost_train,response=9,method='cm',
                      alpha=1,nfolds=10,grouped=TRUE)
pred.cm.lasso<-predictsym.glm(res.cm.lasso,response=9,int_prost_test,method='cm')
plot(res.cm.lasso)
plot(res.cm.lasso$glmnet.fit, "lambda", label=TRUE)
RMSE.L(sym.var(int_prost_test,9),pred.cm.lasso)
RMSE.U(sym.var(int_prost_test,9),pred.cm.lasso)
R2.L(sym.var(int_prost_test,9),pred.cm.lasso)
R2.U(sym.var(int_prost_test,9),pred.cm.lasso)
deter.coefficient(sym.var(int_prost_test,9),pred.cm.lasso)
```

## RIDGE

```{r,fig.align='center'}
data(int_prost_train)
data(int_prost_test)
res.cm.ridge<-sym.glm(sym.data=int_prost_train,response=9,method='cm',
                      alpha=0,nfolds=10,grouped=TRUE)
pred.cm.ridge<-predictsym.glm(res.cm.ridge,response=9,int_prost_test,method='cm')
plot(res.cm.ridge)
plot(res.cm.ridge$glmnet.fit, "lambda", label=TRUE)
RMSE.L(sym.var(int_prost_test,9),pred.cm.ridge)
RMSE.U(sym.var(int_prost_test,9),pred.cm.ridge)
R2.L(sym.var(int_prost_test,9),pred.cm.ridge)
R2.U(sym.var(int_prost_test,9),pred.cm.ridge)
deter.coefficient(sym.var(int_prost_test,9),pred.cm.ridge)
```

# Symbolic Correspondance Analysis

```{r,fig.align='center'}
data(ex_cfa1) 
res <- sym.cfa(ex_cfa1)
cfa.scatterplot(sym.var(res,1), sym.var(res,2), num.gr1=ex_cfa1$N,
                        labels=TRUE,col='red',main='CFA')
```

